---
title: "Differential plasma miRNA profiles during the peri-implantation period in patients with different reproductive outcomes"
author:
  - Zeng, Hong
documentclass: ctexart
geometry: "left=0.5cm,right=0.5cm,top=1.0cm,bottom=1.0cm"
output:
  rticles::ctex:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 5
---
# Color set
```{r}
library(RColorBrewer)
colorset <- brewer.pal(8, "Accent")
# barplot(rep(1,8),col = brewer.pal(8, "Accent"))
```


# Including all Samples
## load data and data clean
```{r data preparation}
load(file = "upload_all_samples.Rda")#load data
keep<-rowSums(expr_tpm>5)>=5#filter criteria
expr_count_filter<-expr_count[keep, ]#filtered count data
expr_tpm_filter<-expr_tpm[keep, ]#filtered TPM data
rm(keep)
```

## PCA analysis
```{r}
pca_all<-prcomp(t(expr_tpm_filter),scale=T)#PCA
df_pca_all<-data.frame(pca_all$x)#Extraction of principal components
library(tidyverse)
df_pca_all <- df_pca_all %>% rownames_to_column(var = "sample_id") %>% left_join(target[,c("sample_id", "group", "ETday")], by="sample_id") %>% column_to_rownames(var = "sample_id")#Add grouping information
head(df_pca_all,3) 
```

## PCA 3D plot
```{r PCA 3D plot}
library(scatterplot3d)
mycolors_pca <- ifelse(df_pca_all$group=="CP", colorset[1], ifelse(df_pca_all$group=="BP", colorset[2], colorset[3]))

pchs <- ifelse(df_pca_all$ETday=="0", 15, ifelse(df_pca_all$ETday=="11", 16, 17))

scatterplot3d(df_pca_all[,1:3],
              color = mycolors_pca,
              pch=pchs,
              main= "",
              grid=T, 
              box=FALSE, 
              type = "h",
              cex.symbols=0.5)
# Add legends
legend("top", legend = c("NP", "BPL", "CP"),
      col =  colorset[c(3,2,1)],  pch =c(NA,NA,NA), inset = 0.06, horiz=T, xpd = T, bty = "n",lwd = 1, lty = c(1,1,1))
legend("top", legend = c("ET0", "ET11", "ET14"),
      col =  c("black", "black", "black"),  pch =c(15,16,17), horiz=T, xpd = T, bty = "n")
```


# Including Samples at ET0
## Load data
```{r data and filter, message=FALSE}
load("upload_samples_ET0.Rda")#load data
keep <- rowSums(expr_tpm_ET0 > 5) >= 5#filter criteria
expr_count_ET0_filter <- expr_count_ET0[keep, ]
expr_tpm_ET0_filter <- expr_tpm_ET0[keep, ]
rm(keep)
```

## PCA analysis
```{r PCA analysis}
pca_ET0<-prcomp(t(expr_tpm_ET0_filter),scale=T)#PCA
df_pca_ET0<-data.frame(pca_ET0$x)#Extraction of principal components
library(tidyverse)
df_pca_ET0 <- df_pca_ET0 %>% rownames_to_column(var = "sample_id") %>% left_join(target_ET0[,c("sample_id", "group")], by="sample_id") %>% column_to_rownames(var = "sample_id")
head(df_pca_ET0,3)
```

## PCA 3D plot
```{r 3D PCA plot}
library(scatterplot3d)
mycolors_ET0 <- ifelse(df_pca_ET0$group=="CP", colorset[1], ifelse(df_pca_ET0$group=="BP", colorset[2], colorset[3]))
scatterplot3d(df_pca_ET0[,1:3],
              color = mycolors_ET0,
              pch=15,
              main= "",
              grid=T, 
              box=FALSE, 
              type = "h")
legend("top", legend = c("NP", "BPL", "CP"),
      col =  colorset[c(3,2,1)],  pch =15, horiz=T, xpd = T, bty = "n",lwd = 1, lty = c(1,1,1))
```


## DEmiRs analysis
```{r DESeq2 analysis, echo=FALSE, message=FALSE, warning=FALSE}
library(DESeq2)
dds_ET0_adjust <- DESeqDataSetFromMatrix(
  countData = expr_count_ET0_filter,
  colData = target_ET0,
  design = ~ embryo_no+embryo_type+group)#adjust for the number of embryos and the type of embryos
dds_ET0_adjust <- DESeq(dds_ET0_adjust, parallel = T) #standardization and DEG analysis
resultsNames(dds_ET0_adjust)
```

```{r DEmiR BPL vs CP at ET0}
library(tidyverse)
res_BPvsCP_ET0 <- results(dds_ET0_adjust, contrast = c("group", "BP", "CP" ),cooksCutoff = F, independentFiltering = F) #BPL vs CP

DEG_BPvsCP_ET0 <- as.data.frame(res_BPvsCP_ET0) %>% 
  rownames_to_column("miRNA_id") %>% 
  arrange(pvalue) %>%
  mutate(Regulation = ifelse(pvalue < 0.05 & abs(log2FoldChange) > 1, ifelse(log2FoldChange < 0,"DOWN", "UP"), "NS"))
# write.csv(DEG_BPvsCP_ET0, file = "Stable_DEG_ET0_BPLvsCP.csv", row.names = F)
DEG_sig_BPvsCP_ET0 <- as.data.frame(res_BPvsCP_ET0) %>% 
  rownames_to_column("miRNA_id") %>% 
  dplyr::filter(abs(log2FoldChange) > 1, pvalue < 0.05) %>%
  arrange(pvalue) %>%
  mutate(Regulation = ifelse(log2FoldChange < 0, "DOWN", "UP"))
```

```{r DEmiR BPL vs NP at ET0}
res_BPvsNP_ET0 <- results(dds_ET0_adjust, contrast = c("group", "BP", "NP" ),cooksCutoff = F, independentFiltering = F) #BPL vs NP

DEG_BPvsNP_ET0 <- as.data.frame(res_BPvsNP_ET0) %>% 
  rownames_to_column("miRNA_id") %>% 
  arrange(pvalue) %>%
  mutate(Regulation = ifelse(pvalue < 0.05 & abs(log2FoldChange) > 1, ifelse(log2FoldChange < 0,"DOWN", "UP"), "NS"))
# write.csv(DEG_BPvsNP_ET0, file = "Stable_DEG_ET0_BPLvsNP.csv", row.names = F)

DEG_sig_BPvsNP_ET0 <- as.data.frame(res_BPvsNP_ET0) %>% 
  rownames_to_column("miRNA_id") %>% 
  dplyr::filter(abs(log2FoldChange) > 1, pvalue < 0.05) %>%
  arrange(pvalue) %>% 
  mutate(Regulation = ifelse(log2FoldChange < 0, "DOWN", "UP"))
```

```{r intersection of two comparisions}
BPL_miRs_ET0 <- intersect(DEG_sig_BPvsCP_ET0$miRNA_id, DEG_sig_BPvsNP_ET0$miRNA_id)
BPL_miRs_ET0
# write.csv(BPL_miRs_ET0, file = "BPL_miRs_ET0.csv", row.names = F)
```

```{r DEmiR NP vs CP at ET0}
res_NPvsCP_adjust <- results(dds_ET0_adjust, contrast = c("group", "NP", "CP" ),cooksCutoff = F, independentFiltering = F) #提取结果
DEG_NPvsCP_ET0 <- as.data.frame(res_NPvsCP_adjust) %>% 
  rownames_to_column("miRNA_id") %>% 
  arrange(pvalue) %>%
  mutate(Regulation = ifelse(pvalue < 0.05 & abs(log2FoldChange) > 1, ifelse(log2FoldChange < 0,"DOWN", "UP"), "NS"))
# write.csv(DEG_NPvsCP_ET0, file = "Stable_DEG_ET0_NPvsCP.csv", row.names = F)

DEG_sig_NPvsCP_ET0 <- as.data.frame(res_NPvsCP_adjust) %>% 
  rownames_to_column("miRNA_id") %>% 
  dplyr::filter(abs(log2FoldChange) > 1, pvalue < 0.05) %>% 
  arrange(pvalue) %>% 
  mutate(Regulation = ifelse(log2FoldChange < 0, "DOWN", "UP")) 
```

 
# Including Samples at ET11
## load data
```{r data and filter}
load(file = "upload_samples_ET11.Rda")# loading data
keep <- rowSums(expr_tpm_ET11 > 5) >= 5#filter
expr_count_ET11_filter <- expr_count_ET11[keep, ]
expr_tpm_ET11_filter <- expr_tpm_ET11[keep, ]
rm(keep)
```

## PCA analysis
```{r PCA analysis}
pca_ET11<-prcomp(t(expr_tpm_ET11_filter),scale=T)
df_pca_ET11<-data.frame(pca_ET11$x)
library(tidyverse)
df_pca_ET11 <- df_pca_ET11 %>% rownames_to_column(var = "sample_id") %>% left_join(target_ET11[,c("sample_id", "group")], by="sample_id") %>% column_to_rownames(var = "sample_id")
head(df_pca_ET11,3)
```

```{r 3D PCA plot}
library(scatterplot3d)
mycolors_ET11 <- ifelse(df_pca_ET11$group=="CP", colorset[1], ifelse(df_pca_ET11$group=="BP", colorset[2], colorset[3]))
scatterplot3d(df_pca_ET11[,1:3],
              color = mycolors_ET11,
              pch=16,
              main= "",
              grid=T, 
              box=FALSE, 
              type = "h")
legend("top", legend = c("NP", "BPL", "CP"),
      col =  colorset[c(3,2,1)],  pch =16, horiz=T, xpd = T, bty = "n",lwd = 1, lty = c(1,1,1))
```

## DEmiRs analysis
```{r DESeq2 analysis, echo=FALSE, message=FALSE, warning=FALSE}
library(DESeq2)
identical(colnames(expr_count_ET11_filter), target_ET11$sample_id)
dds_ET11_adjust <- DESeqDataSetFromMatrix(countData = expr_count_ET11_filter,
                               colData = target_ET11,
                               design = ~ embryo_no+embryo_type+group)#adjust for the number of embryos and the type of embryos
dds_ET11_adjust <- DESeq(dds_ET11_adjust, parallel = T) #standardization and DEG analysis
resultsNames(dds_ET11_adjust)
```

```{r DEmiR BPL vs CP at ET11}
library(tidyverse)
res_BPvsCP_ET11 <- results(dds_ET11_adjust, contrast = c("group", "BP", "CP" ),cooksCutoff = F, independentFiltering = F) #BPL vs CP

DEG_BPvsCP_ET11 <- as.data.frame(res_BPvsCP_ET11) %>% 
  rownames_to_column("miRNA_id") %>% 
  arrange(pvalue) %>%
  mutate(Regulation = ifelse(pvalue < 0.05 & abs(log2FoldChange) > 1, ifelse(log2FoldChange < 0,"DOWN", "UP"), "NS"))
# write.csv(DEG_BPvsCP_ET11, file = "Stable_DEG_ET11_BPLvsCP.csv", row.names = F)
DEG_sig_BPvsCP_ET11 <- as.data.frame(res_BPvsCP_ET11) %>% 
  rownames_to_column("miRNA_id") %>% 
  dplyr::filter(abs(log2FoldChange) > 1, pvalue < 0.05) %>%
  arrange(pvalue) %>%
  mutate(Regulation = ifelse(log2FoldChange < 0, "DOWN", "UP"))
```

```{r DEmiR BPL vs NP at ET11}
res_BPvsNP_ET11 <- results(dds_ET11_adjust, contrast = c("group", "BP", "NP" ),cooksCutoff = F, independentFiltering = F)#BPL vs NP
DEG_BPvsNP_ET11 <- as.data.frame(res_BPvsNP_ET11) %>% 
  rownames_to_column("miRNA_id") %>% 
  arrange(pvalue) %>%
  mutate(Regulation = ifelse(pvalue < 0.05 & abs(log2FoldChange) > 1, ifelse(log2FoldChange < 0,"DOWN", "UP"), "NS"))
# write.csv(DEG_BPvsNP_ET11, file = "Stable_DEG_ET11_BPLvsNP.csv", row.names = F)
DEG_sig_BPvsNP_ET11 <- as.data.frame(res_BPvsNP_ET11) %>% 
  rownames_to_column("miRNA_id") %>% 
  dplyr::filter(abs(log2FoldChange) > 1, pvalue < 0.05) %>%
  arrange(pvalue) %>% 
  mutate(Regulation = ifelse(log2FoldChange < 0, "DOWN", "UP"))
```

```{r intersection of two comparisons}
BPL_miRs_ET11 <- intersect(DEG_sig_BPvsCP_ET11$miRNA_id, DEG_sig_BPvsNP_ET11$miRNA_id)
BPL_miRs_ET11
# write.csv(BPL_miRs_ET11, "BPL_miRs_ET11.csv", row.names = F)
```

```{r DEmiR NP vs CP at ET11}
res_NPvsCP_ET11 <- results(dds_ET11_adjust, contrast = c("group", "NP", "CP" ),cooksCutoff = F, independentFiltering = F)#NP vs CP
DEG_NPvsCP_ET11 <- as.data.frame(res_NPvsCP_ET11) %>% 
  rownames_to_column("miRNA_id") %>% 
  arrange(pvalue) %>%
  mutate(Regulation = ifelse(pvalue < 0.05 & abs(log2FoldChange) > 1, ifelse(log2FoldChange < 0,"DOWN", "UP"), "NS"))
# write.csv(DEG_NPvsCP_ET11, file = "Stable_DEG_ET11_NPvsCP.csv", row.names = F)
DEG_sig_NPvsCP_ET11 <- as.data.frame(res_NPvsCP_ET11) %>% 
  rownames_to_column("miRNA_id") %>% 
  dplyr::filter(abs(log2FoldChange) > 1, pvalue < 0.05) %>%
  arrange(pvalue) %>% 
  mutate(Regulation = ifelse(log2FoldChange < 0, "DOWN", "UP"))
```

# Including Samples at ET14
## load data
```{r data and filter}
load(file = "upload_samples_ET14.Rda")
keep <- rowSums(expr_tpm_ET14>5)>=5
expr_count_ET14_filter <- expr_count_ET14[keep, ]
expr_tpm_ET14_filter <- expr_tpm_ET14[keep, ]
rm(keep)
```

## PCA analysis
```{r PCA analysis}
pca_ET14<-prcomp(t(expr_tpm_ET14_filter),scale=T)
df_pca_ET14<-data.frame(pca_ET14$x)
library(tidyverse)
df_pca_ET14 <- df_pca_ET14 %>% rownames_to_column(var = "sample_id") %>% left_join(target_ET14[,c("sample_id", "group")], by="sample_id") %>% column_to_rownames(var = "sample_id")
head(df_pca_ET14,3)
```

```{r 3D PCA plot}
library(scatterplot3d)
mycolors_ET14 <- ifelse(df_pca_ET14$group=="CP", colorset[1], colorset[2])
scatterplot3d(df_pca_ET14[,1:3],
              color = mycolors_ET14,
              pch=17,
              main= "",
              grid=T, 
              box=FALSE, 
              type = "h")
legend("top", legend = c("BPL", "CP"),
      col =  colorset[c(2,1)],  pch =17, horiz=T, xpd = T, bty = "n",lwd = 1, lty = c(1,1,1))
```

## DEmiRs analysis
```{r DESeq2 analysis, echo=FALSE}
library(DESeq2)
identical(colnames(expr_count_ET14_filter), target_ET14$sample_id)
dds_ET14_adjust <- DESeqDataSetFromMatrix(countData = expr_count_ET14_filter, colData = target_ET14,design = ~ embryo_no+embryo_type+group)#adjust for the number of embryos and the type of embryos
dds_ET14_adjust <- DESeq(dds_ET14_adjust, parallel = T) #standardization and DEG analysis
resultsNames(dds_ET14_adjust)
```

```{r DEmiR BPL vs CP at ET14}
library(tidyverse)
res_BPvsCP_ET14 <- results(dds_ET14_adjust, contrast = c("group", "BP", "CP" ),cooksCutoff = F, independentFiltering = F) #BPL vs CP
DEG_BPvsCP_ET14 <- as.data.frame(res_BPvsCP_ET14) %>% 
  rownames_to_column("miRNA_id") %>% 
  arrange(pvalue) %>%
  mutate(Regulation = ifelse(pvalue < 0.05 & abs(log2FoldChange) > 1,ifelse(log2FoldChange < 0,"DOWN", "UP"), "NS"))
# write.csv(DEG_BPvsCP_ET14, file = "Stable_DEG_ET14_BPLvsCP.csv", row.names = F)
DEG_sig_BPvsCP_ET14 <- as.data.frame(res_BPvsCP_ET14) %>% 
  rownames_to_column("miRNA_id") %>% 
  dplyr::filter(abs(log2FoldChange) > 1, pvalue < 0.05) %>%
  arrange(pvalue) %>%
  mutate(Regulation = ifelse(log2FoldChange < 0, "DOWN", "UP"))

BPL_miRs_ET14 <- DEG_sig_BPvsCP_ET14$miRNA_id
# write.csv(BPL_miRs_ET14, file = "BPL_miRs_ET14.csv", row.names = F)
```

# BPL indicative miRNAs at ET0 and ET11
```{r}
BPL_indicators <- union(BPL_miRs_ET0, BPL_miRs_ET11)
BPL_indicators
# write.csv(BPL_indicators, file = "BPL_indicators.csv", row.names = F)
```


# Heatmap
## ET0
```{r pheatmap at ET0}
library(pheatmap)
library(tidyverse)
heatmap_miRs_ET0 <- rbind(DEG_sig_BPvsCP_ET0[c(1:10),], DEG_sig_BPvsNP_ET0[c(1:10),], DEG_sig_NPvsCP_ET0[c(1:10),])
heatmap_miRs_ET0<-heatmap_miRs_ET0[!duplicated(heatmap_miRs_ET0$miRNA_id),]
heatmap_df_ET0<-expr_tpm_ET0_filter[heatmap_miRs_ET0$miRNA_id, ]

annotation_col_ET0 <- target_ET0 %>%
  dplyr::select(sample_id, group) %>% 
  arrange(group) %>% 
  column_to_rownames("sample_id")

heatmap_df_ET0<-heatmap_df_ET0[,rownames(annotation_col_ET0)]
heatmap_df_ET0 <- heatmap_df_ET0 %>% rownames_to_column(var = "miRNA_id") 
heatmap_df_ET0$miRNA_id <- gsub(heatmap_df_ET0$miRNA_id, pattern = "hsa-", replacement = "")
heatmap_df_ET0 <- heatmap_df_ET0 %>% column_to_rownames(var = "miRNA_id")
```

```{r}
pdf("Heatmap_ET0.pdf", 7, 8, family = "Times")
heatmap_ET0 <- pheatmap(heatmap_df_ET0, 
         scale = "row",
         color = colorRampPalette(c("navy", "white", "#F0027F"))(100),
         border_color = NA,
         main = "ET0", 
         cluster_rows = F,
         cluster_cols = F,
         annotation_col = annotation_col_ET0,
         annotation_colors = list(group=c(CP=colorset[1], BP=colorset[2], NP=colorset[3])),
         annotation_legend = T,
         # treeheight_row = 20,
         # treeheight_col = 15,
         fontsize = 7,
         fontsize_row = 7,
         show_colnames = F,
         legend = F
         # display_numbers = TRUE
         )
ggplot2::ggsave(filename = "heatmap_ET0.pdf", plot = heatmap_ET0, device = "pdf", width = 8, height = 6, units = "cm", dpi = 300, family="Times")
```

## ET11
```{r pheatmap at ET11}
library(pheatmap)
library(tidyverse)
heatmap_miRs_ET11 <- rbind(DEG_sig_BPvsCP_ET11[c(1:10),], DEG_sig_BPvsNP_ET11[c(1:10),], DEG_sig_NPvsCP_ET11[c(1:10),])
heatmap_miRs_ET11<-heatmap_miRs_ET11[!duplicated(heatmap_miRs_ET11$miRNA_id),]
heatmap_df_ET11<-expr_tpm_ET11_filter[heatmap_miRs_ET11$miRNA_id, ]
annotation_col_ET11 <- target_ET11 %>%
  dplyr::select(sample_id, group) %>% 
  arrange(group) %>% 
  column_to_rownames("sample_id")
heatmap_df_ET11<-heatmap_df_ET11[ , rownames(annotation_col_ET11)]
heatmap_df_ET11<-heatmap_df_ET11 %>% rownames_to_column(var = "miRNA_id") 
heatmap_df_ET11$miRNA_id <- gsub(heatmap_df_ET11$miRNA_id, pattern = "hsa-", replacement = "")
heatmap_df_ET11 <- heatmap_df_ET11 %>% column_to_rownames(var = "miRNA_id")
```

```{r}
heatmap_ET11 <- pheatmap(heatmap_df_ET11, 
         scale = "row",
         color = colorRampPalette(c("navy", "white", "#F0027F"))(100),
         border_color = NA,
         main = "ET11", 
         cluster_rows = F,
         cluster_cols = F,
         annotation_col = annotation_col_ET11,
         annotation_colors = list(group=c(CP=colorset[1], BP=colorset[2], NP=colorset[3])),
         annotation_legend = T,
         # treeheight_row = 20,
         # treeheight_col = 15,
         fontsize = 7,
         fontsize_row = 7,
         show_colnames = F,
         legend = F
         # display_numbers = TRUE
         )
ggplot2::ggsave(filename = "heatmap_ET11.pdf", plot = heatmap_ET11, device = "pdf", width = 8, height = 6, units = "cm", dpi = 300, family="Times")
```

## ET14
```{r pheatmap at ET14}
library(pheatmap)
library(tidyverse)
heatmap_miRs_ET14 <- DEG_sig_BPvsCP_ET14[c(1:25),]
heatmap_df_ET14<-expr_tpm_ET14_filter[heatmap_miRs_ET14$miRNA_id, ]
annotation_col_ET14 <- target_ET14 %>%
  dplyr::select(sample_id, group) %>% 
  arrange(group) %>% 
  column_to_rownames("sample_id")
heatmap_df_ET14<-heatmap_df_ET14[ , rownames(annotation_col_ET14)]
heatmap_df_ET14 <- heatmap_df_ET14 %>% rownames_to_column(var = "miRNA_id") 
heatmap_df_ET14$miRNA_id <- gsub(heatmap_df_ET14$miRNA_id, pattern = "hsa-", replacement = "")
heatmap_df_ET14 <- heatmap_df_ET14 %>% column_to_rownames(var = "miRNA_id")
```

```{r}
heatmap_ET14 <- pheatmap(heatmap_df_ET14, 
         scale = "row",
         color = colorRampPalette(c("navy", "white", "#F0027F"))(100),
         border_color = NA,
         main = "ET14",
         cluster_rows = F,
         cluster_cols = F,
         annotation_col = annotation_col_ET14,
         annotation_colors = list(group=c(CP=colorset[1], BP=colorset[2])),
         annotation_legend = T,
         # treeheight_row = 20,
         # treeheight_col = 15,
         fontsize_row = 7,
         fontsize = 7,
         show_colnames = F,
         legend = T
         # display_numbers = TRUE
         )
ggplot2::ggsave(filename = "heatmap_ET14.pdf", plot = heatmap_ET14, device = "pdf", width = 8, height = 6, units = "cm", dpi = 300, family="Times")
```


# Venn plot
## ET0
```{r}
library(VennDiagram)
venn_ET0 <- venn.diagram(x=list("BPL vs CP" = DEG_sig_BPvsCP_ET0$miRNA_id, 
                      "BPL vs NP" = DEG_sig_BPvsNP_ET0$miRNA_id),
             filename = "VennPlot_ET0.tiff",
             fill = colorset[c(4,5)],
             scaled = F,
             cex = 2.5,
             cat.cex = 2.5,
             main = "",
             # main.cex = 2,
             imagetype= "tiff",
             margin=0.2,
             alpha=0.5,
             col = c(NA, NA),
             cat.fontfamily="Times New Roman",
             cat.dist=-0.1
             # cat.fontface="bold"
             )
```

## ET11
```{r}
venn_ET11 <- venn.diagram(x=list("BPL vs CP" = DEG_sig_BPvsCP_ET11$miRNA_id, 
                      "BPL vs NP" = DEG_sig_BPvsNP_ET11$miRNA_id),
             filename = "VennPlot_ET11.tiff",
             fill = colorset[c(6,7)],
             scaled = F,
             cex = 2.5,
             cat.cex = 2.5,
             main = "",
             # main.cex = 2,
             imagetype= "tiff",
             margin=0.2,
             alpha=0.5,
             col = c(NA, NA),
             cat.fontfamily="Times New Roman",
             cat.dist=-0.1
             # cat.fontface="bold"
             )
```


```{r}
venn_BPL <- venn.diagram(x=list("ET0"=BPL_miRs_ET0, "ET11"=BPL_miRs_ET11),
             filename = "VennPlot_BPL_indicators.tiff",
             fill = colorset[c(1,2)],
             scaled = F,
             cex = 2.5,
             cat.cex = 2.5,
             main = "",
             # main.cex = 2,
             imagetype= "tiff",
             margin=0.2,
             alpha=0.5,
             col = NA,
             cat.dist=-0.1,
             fontfamily="serif",
             cat.pos=0,
             # rotation.degree=180,
             cat.fontfamily="Times New Roman",
             # cat.fontface="bold"
             sep.dist=0
             )
```

# miRNA target mRNAs
```{r}
BPL_miRNAs <- c(BPL_miRs_ET0, BPL_miRs_ET11)
library(readxl)
df_targets <- read_excel("miRTarBase_MTI.xlsx")#miRTarBase_MTI.xlsx file is downloaded from miRTarBase (https://mirtarbase.cuhk.edu.cn/~miRTarBase/miRTarBase_2022/php/index.php)
df_targets %>% group_by(miRNA) %>% summarize(count=n())
df_targets <- df_targets %>% dplyr::select(miRNA, `Target Gene`) %>% filter(miRNA %in% BPL_miRNAs)
genes_enri <- df_targets[!duplicated(df_targets$`Target Gene`), 2]
BPL_mRNA <- read.csv("/Users/minizenghong/R_DataFile/miRNA_project/df_BPLindicator.csv", header = F)
mRNAs <- intersect(BPL_mRNA$V1, genes_enri$`Target Gene`)
```

# Enrichment analysis
```{r}
library(clusterProfiler)
library(org.Hs.eg.db)
library(tidyverse)
gene.df <- bitr(genes_enri$`Target Gene`,
                fromType = "SYMBOL",
                toType = c("ENSEMBL", "ENTREZID"),
                OrgDb = org.Hs.eg.db)
gene <- gene.df$ENTREZID

#GO enrichment analysis
ego <- enrichGO(gene = gene,
                ont = "ALL",
                OrgDb = org.Hs.eg.db, 
                pvalueCutoff =0.05, 
                qvalueCutoff =0.05,
                readable = TRUE)

ego_MF <- enrichGO(gene = gene,
                OrgDb = org.Hs.eg.db, 
                pvalueCutoff =0.05, 
                qvalueCutoff =0.05,
                ont = "MF",
                readable = TRUE)

ego_BP <- enrichGO(gene = gene,
                OrgDb = org.Hs.eg.db, 
                pvalueCutoff =0.05, 
                qvalueCutoff =0.05,
                ont = "BP",
                readable = TRUE)

ego_CC <- enrichGO(gene = gene,
                OrgDb = org.Hs.eg.db, 
                pvalueCutoff =0.05, 
                qvalueCutoff =0.05,
                ont = "CC",
                readable = TRUE)

#KEGG enrichment analysis
kk <- enrichKEGG(gene = gene, 
                 organism = "hsa", 
                 pvalueCutoff =0.05, 
                 qvalueCutoff = 0.2)

# save(ego,ego_BP,ego_CC,ego_MF, kk, file = "clusterProfiler_result.Rdata")
```

```{r}
## Visulaization of enrichment analysis
load(file = "clusterProfiler_result.Rda")

pdf("GO_dotplot_BP.pdf", 7, 8, family = "Times")
dotplot(ego_BP, showCategory = 15,color = "pvalue", font.size = 16)
dev.off()

pdf("GO_dotplot_MF.pdf", 7, 8, family = "Times")
dotplot(ego_MF, showCategory = 15,color = "pvalue", font.size = 16)
dev.off()

pdf("GO_dotplot_CC.pdf", 7, 8, family = "Times")
dotplot(ego_CC, showCategory = 15,color = "pvalue", font.size = 16)
dev.off()

pdf("KEGG_dotplot.pdf", 7, 8, family = "Times")
dotplot(kk, showCategory = 15,color = "pvalue", font.size = 16)
dev.off()
```

# Sankey diagram of miRNA-mRNA interactions
```{r}
library(tidyverse)
df_Sankey <- df_targets %>% dplyr::filter(`Target Gene` %in% BPL_mRNA$V1)
df_Sankey <- df_Sankey[!duplicated(df_Sankey),]
names(df_Sankey) <- c("source", "target")
nodes_Sankey <- data.frame(name=c(as.character(df_Sankey$source), as.character(df_Sankey$target)) %>% unique())
df_Sankey$IDsource<-match(df_Sankey$source, nodes_Sankey$name)-1
df_Sankey$IDtarget<-match(df_Sankey$target, nodes_Sankey$name)-1
df_Sankey <- as.data.frame(df_Sankey)
df_Sankey$value <- 1
library(viridisLite)
library(viridis)
library(patchwork)
library(hrbrthemes)
library(circlize)
library(networkD3)
sankeyNetwork(Links = df_Sankey, Nodes = nodes_Sankey,
              Source = "IDsource", Target = "IDtarget",
              Value = "value", NodeID = "name", LinkGroup = 'source',
              sinksRight=FALSE, colourScale=JS("d3.scaleOrdinal(d3.schemeCategory20);"), nodeWidth=40, fontSize=20, nodePadding=20)
```

